### 2013-03-12 ISMB/supplementary_ISMB.tex
remove SSD from supp mat

%!TEX root = main_ISMB.tex
\section{Supplementary data}

\subsection{Benchmark other softwares}
To evaluate the performances of \ourprog, we benchmark a set of classical softwares lacking \GCContent control. Those are \RNAinverse, \INFORNA, \NUPACK and \frankenstein. We present in Fig.~\ref{fig:no_gc_control} the average sequence identity and frequency
for sequences generated them.

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=\textwidth]{Figures/entropy_nogc}
  \caption{The average sequence identity and frequency for softwares without \GCContent control.}
  \label{fig:no_gc_control}
\end{figure*}

\subsection{Benchmark \ourprog+\RNAinverse}
To emphasize the usefulness of processing \ourprog sequences with \RNAinverse, we present the number
of structures for which at least one sequence was generated with the desired MFE in Figure.~\ref{fig:nb_sols}

\begin{figure*}[ht!]
  \centering
  \includegraphics[width=\textwidth]{Figures/struct_solved_vsrnainverse.png}
  \caption{The first row shows the number of structures for which one generated sequence has the 
  structure as MFE when only using \ourprog. The second row shows when we process \ourprog results with \RNAinverse.}
  \label{fig:nb_sols}
\end{figure*}

%\subsection{Benchmark \ourprog vs \RNASSD}
%Using the same dataset of $50$ structures, we generated $100$ samples
%per structure %with \RNAinverse. They yield for most parameters
%a high entropy.
%Table.~\ref{tab:nb_rnainv} contains the number of samples with a given \GCContent generated
%with  \RNAinverse, \INFORNA, \NUPACK and \frankenstein. They have a highly concentrated distribution of \GC content showing one of theirs major deficiencies. We present in Fig.~\ref{fig:gcdist} those distributions.

%\begin{figure}[ht]
%  \begin{center}
%    \includegraphics[width=0.5\textwidth]{Figures/histograme_5_gc_distribution_nornaexinv.png}
%  \end{center}
%  \caption{Overall \GCContent distribution for sequences designed using \RNAinverse, \INFORNA, \NUPACK and \frankenstein folding in the desired structure.}
%  \label{fig:gcdist}
%\end{figure}

%For all structures that have been solved  by the three methods, only \RNASSD, only \texttt{Incarnation} and \texttt{Incarnation} followed by \RNAinverse, we present for different concentration of \GCContent the average sequence identity in Figure.~\ref{fig:identity_50_70_90}
%the average sample entropy and the average base pairs entropy in Figure.~\ref{fig:rnainverse}.
%\begin{figure*}[ht!]
%  \centering
%  \includegraphics[width=\textwidth]{Figures/seq_identity_50_70_90.png}
%  \caption{Sequence identity by average \GCContent of $50$, $70$ and $90\%$ for \ourprog,\ourprog+\RNAinverse and \RNASSD.}
%  \label{fig:identity_50_70_90}
%\end{figure*}
%\begin{table}[h!]
%	\begin{center}
%		\begin{tabular}{|c|ccccc|}
%		\hline
%		Target \GCContent & 10 & 30 & 50 & 70 & 90\\ \hline
%   $\#$\RNAinverse samples& 11 & 437 & 3362 & 1155 & 35\\ \hline
%		\end{tabular}
%	\end{center}
%  \caption{Overall \GCContent distribution for sequences designed using \RNAinverse.}
%	\label{tab:nb_rnainv}
%\end{table}

%\begin{figure*}[ht!]
%	\centering
	%\hspace{-5em}
%	\includegraphics[width=0.5\textwidth]{Figures/RNAinverse_data_100.png}
%	\caption{Entropy and \GCContent  for structures solved by
%	the 3 methods.}
%	\label{fig:rnainverse}
%\end{figure*}
%{\em Récupérer Base-pair entropy}

%The same analysis but only for structures solved by \texttt{Incarnation}
%and \texttt{Incarnation} followed by \RNAinverse, is presented in 
%Fig.~\ref{fig:inc_rnainv}
%
%\begin{figure}
%	\centering
%	\includegraphics{}
%	\caption{Entropy and \texttt{C+G} content for structures solved by
%	the \texttt{Incarnation} and \texttt{Incarnation}$+$\RNAinverse.}
%	\label{fig:inc_rnainv}
%\end{figure}

\subsection{Limited impact on \GC of local-search postprocessing of \ourprog output}
Since local search approaches tend to experience a bias towards \GC{}-rich regions, it could be expected that our glocal approach, by postprocessing unpaired regions using a local search algorithm, would suffer from such a drift.
However, as summarized in Table~\ref{table:impact_on_gc}, we observed that the local search heuristic used to design nucleotides in loop regions has a very limited impact on the \GCContent. For each class of \GCContent, we reported the observed \GCContent in the sequence initially generated by \ourprog, and the observed \GCContent after the \RNAinverse postprocessing (as defined in Section \ref{subsec:glocal_method}). Our results show that the \GCContent is relatively well conserved (less than 6\% variation), with a general tendency of the postprocessing step to bring the \GCContent back to 50\%.

\begin{table}[ht!]
\begin{center}
\resizebox{0.5\textwidth}{!}{
\begin{tabular}{|c|c|c|}
\hline
\multirow{3}{*}{Target \GCContent (\%)}& \multicolumn{2}{c|}{\GCContent (\%) of designed sequences}\\ \cline{2-3}
 & \ourprog & \ourprog + \RNAinverse\\
 & (Global) & ({Glocal}) \\
\hline
10\% & 15\% & 21\% \quad $\nearrow 6\%$\\
30\% & 30\% & 33\% \quad $\nearrow 3\%$\\
50\% & 48\% & 49\% \quad $\nearrow 1\%$\\
70\% & 71\% & 69\% \quad $\searrow 2\%$\\
90\% & 83\% & 78\% \quad $\searrow 5\%$\\
\hline
\end{tabular}
}
\end{center}
\caption{Observed \GCContent of solutions returned by \ourprog (2nd column) and after the application of the local search postprocessing (3rd column).}
\label{table:impact_on_gc}
\end{table}