### 2015-09-28 ISMB/introduction_ISMB.tex
changed energy values, new energy methods, forward

%!TEX root = main_ISMB.tex
\section{Introduction}
\label{sec:introduction}

At the core of the emerging field of synthetic biology resides our capacity to design and re-engineer molecules with target functions. RNA molecules are well tailored for such applications. The ease to synthesize them (they are directly transcribed from DNA) and the broad diversity of catalytic and regulation functions they can perform enable to integrate \textit{de-novo} logic  circuits within living cells \citep{Rodrigo:2012fk} or re-program existing regulation mechanisms \citep{Chang:2012uq}. Future advances and applications of these techniques in gene-therapy studies will strongly rely on efficient computational methods to design and re-engineer RNA molecules.

Most of RNA functions are, at least partially, encoded by the three-dimensional molecular structures, which are themselves primarily determined by the secondary structures. The development of efficient algorithms for designing RNA sequences with pre-defined secondary structures is thus a milestone to enter the synthetic biology era. \RNAinverse pioneered RNA secondary structure design algorithms. It has been developed and distributed with the Vienna RNA package \citep{Hofacker:1994}. However, only posterior experimental studies revealed the potential and practical impact of these techniques. Thereby, during the last 6 years many improvements and variants of \RNAinverse have been proposed. Conceptually, almost all of existing algorithms follow the same approach. First a seed sequence is selected, then a local search strategy is used to mutate the seed and find, in its vicinity, a sequence with desired folding properties. Using this strategy, \INFORNA \citep{Busch:2006uq}, \RNASSD \citep{Aguirre-Hernandez:2007kx} and \NUPACK \citep{Zadeh:2011fk} significantly improved the performance of RNA secondary structure design algorithms. More recent research studies aimed to include more constraints in the selection criteria. \RNAexinv focused on the design of sequences with enhanced thermodynamical and mutational robustness \citep{Avihoo:2011fk}, while \frankenstein enables to design RNA with multiple target structures \citep{Lyngso:2012vn}.

We recently introduced with \RNAensign a novel paradigm for the search strategy of RNA secondary structure design algorithm \citep{Levin:2012kx}. Instead of a local search approach, we proposed a global sampling strategy of the mutational landscape based on the \RNAmutants algorithm~\citep{Waldispuhl2008}. This methodology offered promising performances, but suffered from  prohibitive runtime and memory consumption. Following our work, Garcia-Martin~\emph{et al} proposed \RNAiFOLD~\citep{Garcia-Martin:2013aa}, an alternate methodology that uses constraint programming techniques to prune the mutational landscape. While also suffering from prohibitive running times, it is worth noting that this latter algorithm also proposes a seed-less approach to the RNA secondary structure design problem.

In this paper, we introduce \ourprog, a RNA secondary structure design algorithm that benefits of our recent algorithmic advances \citep{Reinharz:2013aa} to expand our original \RNAensign algorithm \citep{Levin:2012kx}. \ourprog addresses previous limitations of \RNAensign and offers new functionalities. First, while our previous program had a running time complexity of $\mathcal{O}(n^5)$, \ourprog now runs in linear-time and space complexity, allowing it to demonstrate similar speeds as any local search algorithm. Next, \ourprog is \textit{seed-less}. Unlike \RNAensign, it does not require a seed sequence to initiate its search. Finally, \ourprog implements a novel algorithm based on weighted sampling techniques~\citep{Bodini2010} that enables us to control, for the first time, \textit{explicitly} the \GCContent of the solution. This functionality is essential because wild-type sequences within living organisms often present medium or low \GCContent, presumably to offer better transcription rates and/or structural plasticity. Previous programs do not allow to control this parameter and tend to output sequences having high \GCContent{}s~\citep{Lyngso:2012vn}.

We demonstrate the performance of our algorithms on a set of real RNA structures extracted from the \RNASTRAND database \citep{andronescu2008rna}. To complete this study, we develop an hybrid method combining our global sampling approach with local search strategies such as the one implemented in \RNAinverse.  Remarkably, our glocal methodology overcomes both local and global approaches  for sampling sequences with a specific GC content and target structure.